Python as a machine learning language: experiences from the scikit-learn project

### 2013-06-11 paper.tex
trim the weaknesses section a bit
Needs further trimming and merge with implementation section.

The development team has built up sufficient familiarity
with the sparse/array distinction to factor out common patterns
for supporting both sparse and dense input data structures in many estimators.
In other cases, though, sparse matrix support has to implemented
on a per-estimator basis.
This situation is aggravated by the lack of a C API for sparse matrices,
so that Cython modules have to implement their own version
of operations like sparse dot products
that ideally would have been handled by a library.

However, users of scikit-learn cannot always be expected
to have similar familiarity, and in some cases
(e.g., when users want to leverage feature extraction code
outside the context of a \texttt{Pipeline})
they will have to face the \texttt{scipy.sparse} interface.
This is, of course, as much an artifact of the design choice
to not implement separate classes and interfaces for samples/instances
as it a consequence of SciPy's implementation of sparse matrices.

@@

At this point in time, support for multi-core programming in Python,
at least in the CPython reference implementation targeted by scikit-learn,
is downright weak.
The interpreter supports multithreading, but due to the ``Global Interpreter Lock'' (GIL),
no two threads in a process may at the same time execute Python code.
It follows that multithreading is not a viable solution
even two embarrassingly parallel problems,
such as fitting and evaluation multiple models in a grid search procedure.
The workaround is to use Python's multiprocessing module,
which emulates threads using multiple operating system-level processes,
but this in turn leads to high memory use and overhead
because of the copying of training and test sets
into the various processes' address spaces.\footnote{
  The fact that copying occurs at all is an artifact of NumPy's implementation;
  no attempt at implementing a multiprocessing module
  that shares NumPy arrays using copy-on-write semantics or memory mapping
  has so far been completed.}

\subsection{Python/Cython mismatch}

The Python/Cython divide is another source of potential difficulties
both for users and developers/contributors.
While Cython makes writing performance-critical code possible
in a language that is very much like Python,
doing so properly still requires some knowledge of C.
One source of trouble has been the various fixed-size integer types
that are used by the Python and NumPy C runtimes,
whose sizes are practically all platform-dependent.
Choosing the appropriate integer type from among
\texttt{int}, \texttt{size\_t}, \texttt{Py\_ssize\_t} and \texttt{npy\_intp}
can require knowledge of the Python extension API,
the NumPy C API and the C standard.
Neither Cython nor C compilers provide much in the way of static checking
for the right type,
while picking the wrong type may cause segmentation faults
when users try to use arrays that are too large for the type used
and using a dynamic type often leads to unacceptable slowdowns.

>>>>

The development team has built up sufficient familiarity
with the sparse/array distinction to factor out common patterns
for supporting both sparse and dense input data structures in many estimators.
In other cases, though, sparse matrix support has to implemented
on a per-estimator basis.

@@

At this point in time, support for multi-core programming in Python,
at least in the CPython reference implementation targeted by scikit-learn,
is downright weak.
The interpreter supports multithreading, but due to the ``Global Interpreter Lock'' (GIL),
no two threads in a process may at the same time execute Python code.
It follows that multithreading is not a viable solution
even two embarrassingly parallel problems,
such as fitting and evaluation multiple models in a grid search procedure.
The workaround is to use Python's multiprocessing module,
which emulates threads using multiple operating system-level processes,
but this in turn leads to high memory use and overhead
because of the copying of training and test sets
into the various processes' address spaces.

While Cython makes writing performance-critical code possible
in a language that is very much like Python,
doing so properly still requires some knowledge of C.
One source of trouble has been the various fixed-size integer types
that are used by the Python and NumPy C runtimes,
whose sizes are practically all platform-dependent.
